\section{Introduction}



Poetry is one of the natural human methods to communicate by encapsulating the message in beautiful and artistic manner. William Wordsworth said that poetry is "the spontaneous overflow of powerful feelings: it takes its origin from emotion recollected in tranquillity" \cite{wordsworth1990preface}. More non-philosophic and restricted description defined by \citeauthor{manurung2004evolutionary}, poem is a text that is meaningful, grammatically correct, and poetic \cite{manurung2004evolutionary}.


Automatic poetry generation one challenge in artificial intelligence and computational creativity. It is considerably interesting as it works on human creativity, emotional and intelligence domain \cite{colton2012computational}. The difficulties of this task come from various factors. One major issue lies in the poem evaluation. Evaluating poems and creative arts tend to be subjective, thus automated poetry generation faces difficulty obtaining objective result \cite{binsted1996machine}. Another challenge in poetry generation is the massive requirement of the resources. In order to create a poem that satisfies diverse constraints such as phonetic, syntax, grammar or semantics requires wide arrays of resources \cite{manurung2000towards}. 


The objective of this work is to explore further about poetry generation. This work tries to answer the challenge in computational creativity on automatically builds poetry as one of the human artefacts. This work will be focused on one specific genre of a poem: Haiku. Although specialised, we are hoping that this work can be generalised in some aspects for broader and more generic poetry generation. 


Lastly, poetry generation could be used in various creativity related domain, such as helping tool for song composers, video game company, greeting cards industry, or be used by poet themselves. It also could be used to enhance computer to the human interaction system. Personal assistant systems could embed poetry generation to make it capable to talk in poetic manner, thus making it more friendly towards the user. 


The papers are organised as follows. Section 1 delivers the motivation and the objective of research in automated poem generation. The section 2 of this paper explains necessary background information related to this topic. This section also covers recent related works. The section 3 lists some useful resources to aid the research. Section 4 states the specification of this project. Section 5 describes the detailed information of the proposed methodology. Finally, the section 6 is the last section that discuss evaluation method that will be used in this research.

